Embedding Intuitionistic into Classical Logic

### 2023-03-09 LPAR/LPAR-EIICL.tex
comments

In this section we fix notation and recall the semantics for classical and intuitionistic logic. Furthermore we recapitulate some results from~\cite{otten2005clausal}.

@@

\begin{definition}
	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary predicate symbol $R\in\Sigma$ a $n$-ary predicate $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
	m, &\text{if $b=a$,}\\
	v(b), &\text{otherwise.}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$ for each $n$-ary function symbol $f\in\Sigma$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$
		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.
	\end{itemize}
	 Here $a$ is a free variable that does not occur in $\varphi$. $\varphi$ is satisfiable if $\mathcal M, v\models\varphi$ for some $\mathcal M, v$. It is valid if $\mathcal M, v\models\varphi$ for every $\mathcal M, v$. We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
\end{definition}

@@

An important step in the embedding will be the elimination of quantifiers via Herbrandization.
In this process we introduce fresh variables and add additional function symbols to the signature.
A fresh variable is any variable that does not occur in any considered formula.
Whenever we add a function symbol we implicitly extend the signature by a not previously contained symbol.

@@

Finally we note the results from~\cite{otten2005clausal} which are of importance to us.

\begin{lemma}\label{lemma:propositional-normal-form}
	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R, \mathcal X$, with $\mathcal R$ containing  \emph{flat clauses} of the form
	$$\bigwedge_iA_i\to\bigvee_jB_j$$
	and $\mathcal X$ containing \emph{implication clauses} of the form
	$$(A\to B)\to C,$$
	such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. The size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$. In general they will contain additional propositional variables.
\end{lemma}

\begin{lemma}\label{lemma:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary predicate symbol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_jB_j\right),$$
	$\mathcal X$ containing \emph{implication clauses} of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	and $\mathcal Q$ containing \emph{quantification clauses} of the form
	$$\forall\vec x\left((\forall y A)\to B\right),$$
	such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general this formula will contain new function and predicate symbols.
\end{lemma}

Since we could not find a proof of this, one is given in the appendix~\ref{proof:first-order-normal-form}.

@@

We shall now give an overview of our main results and highlight the key arguments. We start with the propositional case. The important insights are the same as in the first-order case, while it is technically simpler and less cluttered.

The idea is to use the Kripke semantics for intuitionistic logic, expressed in classical logic, to perform an embedding of formulas. That is, for each $\varphi$ we give a $\varphi^\#$ that is classically valid iff $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.

The key property that allows us to reduce our encoding to propositional logic, which a priori targets first order logic, is a reduction of counter-models. In particular, given $$\varphi = \left(\bigwedge\mathcal R\wedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form} we establish:

@@

This allows us to embed intuitionistic propositional logic into classical propositional logic by replacing quantification over worlds by their enumeration.

@@

The exact details can be found in Theorem~\ref{thm:prop-translation}. Instead of expressing correctness at each world of the model, we can express that the worlds along some path in the tree are correct. The branching of paths can be expressed through quantification in a quantified boolean formula (QBF) which gives us the following:

\begin{lemma}
	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
\end{lemma}

Again, the details can be found in Section~\ref{sec:qbf}. Next we attempt to transport these results to the first order case. However, there are formulas for which only counter-examples with an infinite Kripke frame exist. Therefore, there is no hope of completely eliminating terms representing worlds from the final embedding. This also means that our language necessarily becomes more complex. Still, a number of reductions is possible.

@@

\begin{definition}
	Let $\varphi$ be a propositional formula. Define $\varphi^{\nu}$ inductively:
	\vspace*{-.1cm}
	\begin{itemize}
		\item $\bot^\nu = \bot$.
		\item $A^{\nu} = A(\nu)$ for every propositional variable $A$.
		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^\nu$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
	\vspace*{-.1cm}
	\[K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))\]
	\vspace*{-.4cm}
	with e.g.
	\vspace*{-.2cm}
	\begin{align*}
		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	$\varphi^{C} = K(\varphi)\to \varphi^{b}$
	where $b$ is a new free variable.
\end{definition}

@@

\section{IPC to CPC}

@@

\begin{theorem}\label{thm:prop-translation}
	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
	with flat $\mathcal R$ and implication $\mathcal S$. Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$. For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $A^\lambda$. Define $R^\#$ as the union over all sets
	$$
		\left\{\bigwedge_iA_i^\lambda\to\bigvee_jB_j^\lambda\:|\:\lambda\in\Lambda\right\}
	$$
	for $\bigwedge_iA_i\to\bigvee_jB_j\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
	$$
		\left\{\left(A^{\lambda\psi} \to A^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ as the union over all sets
	$$
		\left\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	 for each propositional variable $A$ occurring in $\varphi$, encoding persistency. Then
	$$
		\varphi^\# := \left(\bigwedge P^\#\wedge\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\right)\to P^\epsilon
	$$
	is classically valid if and only if $\varphi$ is intuitionistically valid.
\end{theorem}
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

@@

It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}.
The translation from the last section will serve as starting point for developing the translation in this section.
For lack of space, we do not define the semantics of QBF in this paper and refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.

Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$. Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$. Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec X^i\:|\:i < n\}$.
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulas in $\mathcal R$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
	\end{itemize}
	Define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < |\mathcal{X}|$, as well as the special cases
	$$\varphi^Q_{|\mathcal{X}|} = \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
	for leaves, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulas are:
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee\left\{X^n_\psi\:|\:\psi\in\mathcal X\right\}\right)\wedge\left(\bigwedge \left\{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\right\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j\mathcal R\right\}\\
		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
	\end{align*}
\end{definition}

@@

\begin{lemma}\label{proof:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary relation symbol $P$ as well as sets of clauses $\mathcal R$ of the form
	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_j B_j\right)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	called \emph{implication} clauses as well as $\mathcal Q$ of the form
	$$\forall\vec x\left((\forall y A)\to B\right)$$
	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general this formula will contain new function symbols and predicate symbols.
\end{lemma}

>>>>

In this section we fix notation and recall the semantics for classical and intuitionistic logic. 
Furthermore, we recapitulate some results from~\cite{otten2005clausal}.

@@

\begin{definition}
	Let $\Sigma$ be a signature. A $\Sigma$-structure $\mathcal{M}$ consists of a non-empty set $M$, the \emph{domain} of $\mathcal{M}$, and an \emph{interpretation} $I$ that assigns
	\begin{itemize}
		\item to each $n$-ary function symbol $f\in\Sigma$ a $n$-ary function $f^I: M^n\to M$.
		\item to each $n$-ary predicate symbol $R\in\Sigma$ a $n$-ary predicate $R^I\subseteq M^n$.
	\end{itemize}
	A variable assignment $v$ is a function that assigns to each free variable an element $m\in M$. For each free variable $a$ and $m\in M$ we define $$v[m/a](b) = \begin{cases}
	m, &\text{if $b=a$,}\\
	v(b), &\text{otherwise.}
	\end{cases}$$
	Then terms are interpreted as follows:
	\begin{itemize}
		\item $a^{I, v} = v(a)$ for each free variable $a$.
		\item $f(t_1,\dots,t_n)^{I, v} = f^I(t_1^{I, v},\dots, t_n^{I, v})$ for each $n$-ary function symbol $f\in\Sigma$.
	\end{itemize}
	We define a model relation between pairs $\mathcal M, v$ and formulas $\varphi$ as follows
	\begin{itemize}
		\item $\mathcal M, v\not\models\bot$
		\item $\mathcal M, v\models R(t_1,\dots,t_n)$ iff $(t_1^{I, v},\dots,t_n^{I, v})\in R^I$, $\mathcal M, v\models s = t$ iff $s^{I, v} = t^{I, v}$.
		\item $\mathcal M, v\models \varphi\wedge \psi$ iff $\mathcal M, v\models\varphi$ and $\mathcal M, v\models\psi$
		\item $\mathcal M, v\models \varphi\vee\psi$ iff $\mathcal M, v\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models \varphi\to\psi$ iff $\mathcal M, v\not\models\varphi$ or $\mathcal M, v\models\psi$.
		\item $\mathcal M, v\models\exists x\varphi$ iff there exists $m\in M$ such that $\mathcal M, v[m/a]\models\varphi[a/x]$.
		\item $\mathcal M, v\models\forall x\varphi$ iff for all $m\in M$ we have $\mathcal M, v[m/a]\models\varphi[a/x]$.\todo{Hier und im Rest vom paper bin nicht so glücklich über diese Substitution von durch a..}
	\end{itemize}
	 Here $a$ is a free variable that does not occur in $\varphi$. $\varphi$ is satisfiable if $\mathcal M, v\models\varphi$ for some $\mathcal M, v$. It is valid if $\mathcal M, v\models\varphi$ for every $\mathcal M, v$. We denote the set of valid formulas with \emph{CQC} (Classical Quantified Calculus).
\end{definition}

@@

An important step in the embedding will be the elimination of quantifiers via Herbrandization.
In this process we introduce fresh variables and add additional function symbols to the signature.
A fresh variable is any variable that does not occur in any considered formula.
Whenever we add a function symbol we implicitly extend the signature by a fresh symbol.

@@

Finally we recall a result from~\cite{otten2005clausal}, on which we build in this paper.

\begin{lemma}\label{lemma:propositional-normal-form}
	For every propositional formula $\varphi$ there exists an atom $P$ as well as sets of clauses $\mathcal R, \mathcal X$, with $\mathcal R$ containing  \emph{flat clauses} of the form
	$$\bigwedge_iA_i\to\bigvee_jB_j$$
	and $\mathcal X$ containing \emph{implication clauses} of the form
	$$(A\to B)\to C,$$
	such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. The size of $\mathcal R$ and $\mathcal X$ is linear in the size of $\varphi$. In general these formulae will contain additional propositional variables.
\end{lemma}

\begin{lemma}\label{lemma:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary predicate symbol $P$ as well as sets of clauses $\mathcal R,\mathcal X, \mathcal Q$, with $\mathcal R$ containing \emph{flat clauses} of the form
	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_jB_j\right),$$
	$\mathcal X$ containing \emph{implication clauses} of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	and $\mathcal Q$ containing \emph{quantification clauses} of the form
	$$\forall\vec x\left((\forall y A)\to B\right),$$
	such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general these formulae will contain new function and predicate symbols.
\end{lemma}

Since we could not find a proof for the first-order case, we provide one in the appendix~\ref{proof:first-order-normal-form}.

@@

We shall now give an overview of our main results and highlight the key arguments. 
We start with the propositional case, as the important insights are the same as in the first-order case, while it is technically simpler and less cluttered.

The first step is to express the Kripke semantics of intuitionistic logic in classical first-order logic.
That is, for each $\varphi$ we give a $\varphi^\#$ that is classically valid iff $\varphi$ is intuitionistically valid, establishing a dual to the double negation translation.

Then, the key to reducing the first-order logic encoding to a propositional logic encoding, is a small model property that supports the reduction of counter-models.
In particular, given $$\varphi = \left(\bigwedge\mathcal R\wedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form} we establish:

@@

This allows us to embed intuitionistic propositional logic into classical propositional logic by replacing the quantification over all possible worlds by an explicit enumeration:

@@

The details can be found in Section~\ref{sec:ipc-to-cpc}.
Building on the same idea, we can also give a quantified boolean formula (QBF) of polynomial size, where the branches of the tree-model can be enumerated implicitly using (alternating) quantifiers. We obtain the following result:

\begin{lemma}
	For each propositional formula $\varphi$ there is effectively a QBF $\varphi^Q$ \todo{Ich würde die Formeln $\varphi^Q$ hier und in Section 6 aus Symmetriegründen auch $\varphi^\#$ nennen.} with $\mathcal |\varphi^Q|\in\mathcal O(|\mathcal R|\cdot|\mathcal X| + |\mathcal X|^3)$ such that $\varphi$ is intuitionistically invalid if and only if $\varphi^Q$ is a satisfiable QBF.
\end{lemma}

Again, the details can be found in Section~\ref{sec:qbf}. 
Next we attempt to transport these results to the first order case. However, there are formulas for which only counter-examples with an infinite Kripke frame exist. Therefore, there is no hope of completely eliminating terms representing worlds from the final embedding. This also means that our language necessarily becomes more complex. Still, a number of reductions are possible.

@@

\begin{definition}
	Let $\varphi$ be a propositional formula.\todo{genau genommen müsstest du noch sagen, dass $\nu$ eine freie Variable ist. Im letzten Fall der Definition benützen wir dann aber w gleichzeitig als freie und gebundene Variable}
 Define $\varphi^{\nu}$ inductively:
	\vspace*{-.1cm}
	\begin{itemize}
		\item $\bot^\nu = \bot$.
		\item $A^{\nu} = A(\nu)$ for every propositional variable $A$.
		\item $(\varphi\circ\psi)^\nu = \varphi^\nu\circ\psi^\nu$ for $\circ\in\{\wedge, \vee\}$.
		\item $(\varphi\to \psi)^\nu = \forall w(\nu\preceq w\to\varphi^{w}\to\psi^{w})$ where $w$ is some new bound variable.
	\end{itemize}
	Let $K(\varphi)$ encode the theory of Kripke structures, i.e.
	\vspace*{-.1cm}
	\[K(\varphi) = \text{\normalfont PartialOrder}(\preceq)\wedge\forall u\forall w(u\preceq w\to \text{\normalfont Persistent}(u, w))\]
	\vspace*{-.4cm}
	with e.g.
	\vspace*{-.2cm}
	\begin{align*}
		\text{\normalfont PartialOrder}(\preceq) =&\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
		\text{\normalfont Persistent}(u, w)=&\:\bigwedge \{A(u)\to A(w)\:|\: \text{ $A$ occurs in $\varphi$}\}
	\end{align*}
	Then define
	$\varphi^{C} = K(\varphi)\to \varphi^{b}$
	where $b$ is a new free variable.
\end{definition}

@@

\section{IPC to CPC}
\label{sec:ipc-to-cpc}

@@

\begin{theorem}\label{thm:prop-translation}
	Suppose we are given a formula as in Lemma~\ref{lemma:propositional-normal-form}, i.e. of the form
	$$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal S\right)\to P$$
	with flat $\mathcal R$ and implication $\mathcal S$. Let $\Lambda$ denote the set of sequences without repetition over $\mathcal X$. For every propositional variable $A$ and $\lambda\in\Lambda$ consider a new propositional Variable $A^\lambda$. Define $\mathcal R^\#$ as the union over all sets
	$$
		\left\{\bigwedge_iA_i^\lambda\to\bigvee_jB_j^\lambda\:|\:\lambda\in\Lambda\right\}
	$$
	for $\bigwedge_iA_i\to\bigvee_jB_j\in\mathcal R$ and $\mathcal X^\#$ as the union over all sets
	$$
		\left\{\left(A^{\lambda\psi} \to A^{\lambda\psi}\right)\to C^\lambda\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	for $\psi = (A\to B)\to C\in\mathcal X$ as well as $P^\#$ \todo{Hier ist nur P statt mathcal P, was aus Symmetriegründen vielleicht hilfreich wäre, aber P wird auch schon für die finale Präposition verwendet, also wäre ein anderer Buchstabe wohl besser} as the union over all sets
	$$
		\left\{A^\lambda\to A^{\lambda\psi}\:|\:\lambda\in\Lambda, \psi\text{ does not occur in }\lambda\right\}
	$$
	 for each propositional variable $A$ occurring in $\varphi$, encoding persistency. Then
	$$
		\varphi^\# := \left(\bigwedge P^\#\wedge\bigwedge \mathcal R^\#\wedge\bigwedge\mathcal X^\#\right)\to P^\epsilon
	$$
	is classically valid if and only if $\varphi$ is intuitionistically valid.
\end{theorem}
%\begin{theorem}
%	$\bigwedge\mathcal S\to P$ is intuitionistically valid iff $\bigwedge\mathcal S^\#\to P^\emptyset$ is classically valid.
%\end{theorem}

@@

It is known that a polynomial time translation between QBF and IPC must exist, since both problems are PSPACE complete~\cite{garey1979computers,statman1979intuitionistic}.
\todo{Wollen wir noch etwas sagen wie ``However the previous PSPACE result does not support the transfer of countermodels that we develop in this section''?}
For lack of space, we do not give the semantics of QBF in this paper and instead refer the reader to standard sources e.g.~\cite{series/faia/2009-185}.
The translation from the last section will serve as starting point for this section.

Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}. We now given an intuition on how to obtain a QBF of polynomial size that expresses that $\varphi$ has an intuitionistic counter-model: 
Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. 
The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
For every propositional variable $A$ occurring in $\varphi$ and non-negative integer $n$ (representing the distance of a node from the root on a path) consider a new propositional variable $A^n$ and for every formula $\psi\in\mathcal X$ consider a new propositional variable $X_\psi^n$.
Let $\vec A^n$ range over all propositional variables $A^n$ and $\vec X^n$ range over all $X_\psi^n$. 
Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec X^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec X^i$ as bit-vectors, then $\vec X^n$ has exactly one bit set to $1$, indicating the formula $\psi$ at the $n$-th position of the sequence, and it is at a position that is $0$ in all $\vec X^i$ with $i < n$ (i.e., it did not occur in the sequence previously).
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal R}(n)$ encoding that the formulas in $\mathcal R$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal X}(n)$ encoding that the formula in $\mathcal X$ that is represented by $\vec X^{n-1}$ holds.
	\end{itemize}
	Define$$\varphi^Q_{i} = \exists \vec A^i\forall \vec X^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal R}(i)\wedge \text{\normalfont Sat}_{\mathcal X}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < |\mathcal{X}|$, as well as the special cases
	$$\varphi^Q_{|\mathcal{X}|} = \exists \vec A^{|\mathcal{X}|}\left(\text{\normalfont Persistent}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal R}(|\mathcal{X}|)\wedge \text{\normalfont Sat}_{\mathcal X}(|\mathcal{X}|)\right)$$
	for leaves, and $$\varphi^Q = \varphi_0^Q = \exists \vec A^0\forall \vec X^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal R}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulas are:
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee\left\{X^n_\psi\:|\:\psi\in\mathcal X\right\}\right)\wedge\left(\bigwedge \left\{\neg(X^n_{\psi_1}\wedge X^n_{\psi_2})\:|\:\psi_1\neq\psi_2\in\mathcal X\right\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{X^i_\psi\to\neg X^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal X\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\left\{A^{n-1}\to A^n\:|\: A \text{ prop. variable with }A^{n-1}\in\vec A^{n-1}, A^n \in\vec A^{n}\right\}\\
		\text{\normalfont Sat}_{\mathcal R}(n) = &\bigwedge\left\{\bigwedge A^n_i\to\bigvee B^n_j\:|\:\bigwedge A_i\to\bigvee B_j\mathcal R\right\}\\
		\text{\normalfont Sat}_{\mathcal X}(n) = & \bigwedge\left\{X^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal X\right\}
	\end{align*}
\end{definition}

@@

\begin{lemma}\label{proof:first-order-normal-form}
	For every predicate formula $\varphi$ there exists a nullary relation symbol $P$ as well as sets of clauses $\mathcal R$ of the form
	$$\forall \vec x\left(\bigwedge_i A_i\to \bigvee_j B_j\right)$$
	called \emph{flat} clauses and $\mathcal X$ of the form
	$$\forall \vec x\left((A\to B)\to C\right)$$
	called \emph{implication} clauses as well as $\mathcal Q$ of the form
	$$\forall\vec x\left((\forall y A)\to B\right)$$
	called \emph{quantification} clauses such that $\varphi$ is intuitionistically equivalid to
	$$\left(\bigwedge\mathcal R\wedge\bigwedge \mathcal X\wedge\bigwedge\mathcal Q\right)\to P$$where $A_i, B_i, A, B, C$ are atomic. This size of $\mathcal R, \mathcal X, \mathcal Q$ is linear in the size of $\varphi$. In general these formulae will contain new function symbols and predicate symbols.
\end{lemma}
